---
title: "Homework 4"
author: "Uselton, Myranda"
date: "`r date()`"
output: 
  html_document:
    toc: true
  md_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 4

### Question 4 - ISLR

a) about 10%
b) about 1%
c) near 0 - a very small decimal
d) as the number of features p increases, the KNN looks at a smaller percentage of observations. Thus, there are very few observations near the feature's "location."
e) Length of the sides of the hypercube:
```{r}

p <- 1
lambda <- 0.1^(1/p)
lambda

p <- 2
lambda <- 0.1^(1/p)
lambda

p <- 100
lambda <- 0.1^(1/p)
lambda
```
The sides of the cube grow nearly to 1.


### Question 10 - ISLR

a)

```{r}
library(ISLR)
library(ggplot2)
library(dplyr)
library(caret)
library(e1071)
library(MASS)
library(class)

data <- Weekly

head(data)

summary(data)

ggplot(data, aes(Year, Lag1)) + geom_point()

ggplot(data, aes(Year, Lag2)) + geom_point()

ggplot(data, aes(Year, Lag3)) + geom_point()

ggplot(data, aes(Year, Lag4)) + geom_point()

ggplot(data, aes(Year, Lag5)) + geom_point()
```

The Lag variables tend to be more spread out around the year 2000 and condense the following years

b)

```{r}
model_data <- data %>% 
  select(Direction, Lag1, Lag2, Lag3, Lag4, Lag5, Volume)

model_data$Direction <- ifelse(model_data$Direction == 'Down', 0, 1)

model <- lm(Direction ~ ., model_data)

model

```


All of the variables have a negative coefficient except for Lag2.

c)

```{r}

preds <- predict(model)

preds <- ifelse(preds > 0.5, 1, 0)

preds <- as.factor(preds)

model_data$Direction <- as.factor(model_data$Direction)

confusionMatrix(data = preds, reference = model_data$Direction)


```

The model is about 56% accurate.

d)

```{r}
train <- data %>% 
  filter(Year <= 2008)

test <- data %>% 
  filter(Year > 2008)

test$Direction <- ifelse(test$Direction == 'Down', 0, 1)
train$Direction <- ifelse(train$Direction == 'Down', 0, 1)

model <- lm(Direction ~ Lag2, train)

preds <- predict(model, test)

preds <- ifelse(preds > 0.5, 1, 0)

preds <- as.factor(preds)
test$Direction <- as.factor(test$Direction)

confusionMatrix(data = preds, reference = test$Direction)
```

e)

```{r}
model <- lda(Direction ~ Lag2, train)

preds <- predict(model, test)

preds$class <- as.factor(preds$class)
test$Direction <- as.factor(test$Direction)

confusionMatrix(data = preds$class, reference = test$Direction)
```

f)

```{r}
model <- qda(Direction ~ Lag2, train)

preds <- predict(model, test)

preds$class <- as.factor(preds$class)
test$Direction <- as.factor(test$Direction)

confusionMatrix(data = preds$class, reference = test$Direction)
```

g)

```{r}

cl <- train$Direction
preds <- knn(as.data.frame(train$Lag2), as.data.frame(test$Lag2), cl, k = 1)

confusionMatrix(data = preds, reference = test$Direction)

```

h) The most accurate seems to be lda in (e), with accuracy of 0.625

